---
title: "Learning Bayesian Networks from Ordinal Data - Demo"
author: "Xiang Ge Luo"
date: "2021/04/11"
output: 
  md_document:
    variant: markdown_github
---

### 1. Load necessary packages and source code

```{r, message=FALSE}
library(BiDAG)
library(Rgraphviz)
library(pcalg)
library(graph)
library(bnlearn)
library(MXM)
library(sbgcop)
library(infotheo)
library(rpcart)
library(devtools)
source_url("https://raw.githubusercontent.com/cuiruifei/CausalMissingValues/master/R/inferCopulaModel.R")
```

```{r, message=FALSE}
setwd("../R")
# Major file containing the OSEM algorithm
source("ordinalScore.R") 
# Modify some of the existing functions in the BiDAG package to accommodate our user-defined functions
insertSource("spacefns.R",package = "BiDAG")
insertSource("usrscorefns.R",package = "BiDAG")
insertSource("initpar.R",package = "BiDAG")
insertSource("scoreagainstdag.R",package = "BiDAG")
```

### 2. Generate DAGs and ordinal data

```{r generate}
set.seed(222)
N <- 500
n <- 20

# Generate a regular DAG with 20 nodes with 4 number of neighbors
trueDAG <- randDAG(n = n, d = 4, method = "er", wFUN = list(mywFUN))

# Convert the Gaussian dataset into an ordinal dataset
ordinal_data <- generateOrdinal(N, n, trueDAG, exp_levels = 4, concent_param = 2)
ordinal_data_df <- as.data.frame(ordinal_data)
ordinal_data_df[] <- lapply(ordinal_data_df[], as.ordered)
ordinal_levels <- apply(ordinal_data, 2, function(x) length(unique(x)))
```

### 3. Learn the structures using different methods

The following simulations can be repeated many times for different configurations of parameters in order to obtain the ROC curves.

- NPC

```{r NPC, message=FALSE, warning = FALSE}
# NPC algorithm with the G^2 test (significance level: 0.05)
NPCfit <- pc(suffStat = list(dm = ordinal_data, 
                             nlev = apply(ordinal_data, 2, function (x) length(unique(x))),
                             adaptDF = FALSE),
             alpha = 0.05,
             indepTest = disCItest,
             labels = colnames(ordinal_data))
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as(NPCfit,"matrix"), edgemode = "directed")
plot(tempgraph, main= "NPC")
```

```{r}
# Compare the patterns between them
comparePatterns(NPCfit,trueDAG) # hard version
comparePatterns(NPCfit,trueDAG,hardP2P = FALSE) # soft version
```

- OPC (Musella, 2013)

```{r OPC, message = FALSE, warning = FALSE, cache = TRUE}
# OPC algorithm with the ordinal Jonckheere–Terpstra test (significance level: 0.05) (bnlearn implementation)
OPCfit <- amat(pc.stable(ordinal_data_df, alpha = 0.05, test = "jt"))
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as(OPCfit,"matrix"), edgemode = "directed")
plot(tempgraph, main= "OPC")
```

```{r}
# Compare the patterns between them
comparePatterns(OPCfit,trueDAG) # hard version
comparePatterns(OPCfit,trueDAG,hardP2P = FALSE) #soft version
```

- GPC

```{r GPC, warning = FALSE}
# GPC algorithm with the Gaussian test (significance level: 0.05)
GPCfit <- pc(suffStat = list(C = cor(ordinal_data), n = N),
             alpha = 0.05,
             indepTest = gaussCItest,
             labels = colnames(ordinal_data))
# # GPC algorithm (significance level: 0.05) (bnlearn implementation)
# GPCfit <- amat(pc.stable(as.data.frame(ordinal_data), alpha = 0.05, test = "zf"))
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as(GPCfit,"matrix"), edgemode = "directed")
plot(tempgraph, main= "GPC")
```

```{r}
# Compare the patterns between them
comparePatterns(GPCfit,trueDAG) # hard version
comparePatterns(GPCfit,trueDAG,hardP2P = FALSE) # soft version
```

- RPC (Harris and Drton, 2013; Cui et al., 2018)

```{r RPC, cache = TRUE}
corr.rank <- sin(pi/2 * cor(ordinal_data, use = 'pairwise.complete.obs', method = 'kendall'))
RPCfit <- pc(suffStat = list(C = corr.rank, n = N), 
             indepTest = gaussCItest, labels = colnames(ordinal_data), alpha = 0.05, conservative = T)
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as(RPCfit,"matrix"), edgemode = "directed")
plot(tempgraph, main= "RPC")
```

```{r}
# Compare the patterns between them
comparePatterns(RPCfit,trueDAG) # hard version
comparePatterns(RPCfit,trueDAG,hardP2P = FALSE) # soft version
```

- Copula PC (Cui et al., 2016, 2017, 2018)

```{r CPC, cache = TRUE}
# copula object
cop.obj <- inferCopulaModel(ordinal_data, nsamp = 1000, S0 = diag(n)/N, verb = F)
# correlation matrix samples
C_samples <- cop.obj$C.psamp[,, 501:1000]
# average correlation matrix
corr.cop <- apply(C_samples, c(1,2), mean)
# call the PC algorithm for causal discovery
CPCfit <- pc(suffStat = list(C = corr.cop, n = N), 
             indepTest = gaussCItest, labels = colnames(ordinal_data), alpha = 0.05, conservative = T)
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as(CPCfit,"matrix"), edgemode = "directed")
plot(tempgraph, main= "CPC")
```

```{r}
# Compare the patterns between them
comparePatterns(CPCfit,trueDAG) # hard version
comparePatterns(CPCfit,trueDAG,hardP2P = FALSE) # soft version
```

- MMPC (Tsagris et al., 2018)

```{r MMPC, message=FALSE, warning = FALSE, cache = TRUE}
skel <- pc.skel(ordinal_data_df, method = "comb.mm", alpha = 0.05)
MMDAG <- pc.or(skel)$G
MMDAG[MMDAG == 2] <- 1
MMDAG[MMDAG == 3] <- 0
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(MMDAG, edgemode = "directed")
plot(tempgraph, main= "MMPC")
```

```{r}
# Compare the patterns between them
comparePatterns(MMDAG,trueDAG) # hard version
comparePatterns(MMDAG,trueDAG,hardP2P = FALSE) # soft version
```

- BDe (Heckerman and Geiger, 1995)

```{r BDe, results='hide', cache = TRUE}
# hybrid method with the BDe score and the nominal PC output as the initial search space
BDE <- scoreparameters("bdecat",data.frame(ordinal_data),bdecatpar = list(chi = 0.5))
BDEfit <- iterativeMCMC(BDE)
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as.matrix(BDEfit$DAG), edgemode = "directed")
plot(tempgraph, main= "BDe")
```

```{r}
# Compare the patterns between them
comparePatterns(as.matrix(BDEfit$DAG),trueDAG) # hard version
comparePatterns(as.matrix(BDEfit$DAG),trueDAG,hardP2P = FALSE) # soft version
```

- BGe (Heckerman and Geiger, 1995)

```{r BGe, results='hide', cache = TRUE}
# hybrid method with the BGe score and the Gaussian PC output as the initial search space
BGE <- scoreparameters("bge", ordinal_data, bgepar = list(am = 0.5))
BGEfit <- iterativeMCMC(BGE,scoreout = TRUE)
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as.matrix(BGEfit$DAG), edgemode = "directed")
plot(tempgraph, main= "BGe")
```

```{r}
# Compare the patterns between them
comparePatterns(as.matrix(BGEfit$DAG),trueDAG) # hard version
comparePatterns(as.matrix(BGEfit$DAG),trueDAG,hardP2P = FALSE) # soft version
```

- OSEM

```{r OSEM, results='hide', cache = TRUE}
# the OSEM algorithm
OSEMfit <- ordinalStructEM(n, ordinal_data,
                           usrpar = list(penType = "other",
                                         L = 5,
                                         lambda = 3,
                                         preLevels = NULL))
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as.matrix(OSEMfit$DAG), edgemode = "directed")
plot(tempgraph, main= "OSEM")
```

```{r}
# Compare the patterns between them
comparePatterns(as.matrix(OSEMfit$DAG),trueDAG) # hard version
comparePatterns(as.matrix(OSEMfit$DAG),trueDAG,hardP2P = FALSE) # soft version
```

- PCART (Talvitie et al., 2019)

```{r, message=FALSE}
setwd("../R")
insertSource("usrscorefns_pcart.R",package = "BiDAG")
```

```{r pcart, warning=FALSE, cache=TRUE}
pcartparam <- scoreparameters("usr", ordinal_data_df, 
                              usrpar = list(pcart_alpha = 1.5,
                                            pcart_kappa = 0.25,
                                            pctesttype = "bde", 
                                            preLevels = ordinal_levels,
                                            response_type = "CAT"))
# need to set limit to the parent set size due to computational limit
pcartfit <- iterativeMCMC(pcartparam, alpha = 0, plus1it = 10, softlimit = 3, hardlimit = 3)
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as.matrix(pcartfit$DAG), edgemode = "directed")
plot(tempgraph, main= "PCART")
```

```{r}
# Compare the patterns between them
comparePatterns(as.matrix(pcartfit$DAG),trueDAG) # hard version
comparePatterns(as.matrix(pcartfit$DAG),trueDAG,hardP2P = FALSE) # soft version
```

- OPCART (Talvitie et al., 2019)

```{r opcart, warning=FALSE, cache=TRUE}
opcartparam <- scoreparameters("usr", ordinal_data_df, 
                              usrpar = list(pcart_alpha = 1.5,
                                            pcart_kappa = 0.25,
                                            pctesttype = "bde", 
                                            preLevels = ordinal_levels,
                                            response_type = "ORD"))
# need to set limit to the parent set size due to computational limit
opcartfit <- iterativeMCMC(opcartparam, alpha = 0, plus1it = 10, softlimit = 3, hardlimit = 3)
```

```{r, echo = FALSE}
par(mfrow = c(1,2))
plot(trueDAG, main = "True DAG")
tempgraph <- graphAM(as.matrix(opcartfit$DAG), edgemode = "directed")
plot(tempgraph, main= "OPCART")
```

```{r}
# Compare the patterns between them
comparePatterns(as.matrix(opcartfit$DAG),trueDAG) # hard version
comparePatterns(as.matrix(opcartfit$DAG),trueDAG,hardP2P = FALSE) # soft version
```



